SVA - DEA benchmark

### 2020-05-04 svadea/results/results.Rmd
added sva-dea results

* **ipsc**: 10 vs 10 random samples from the GSE79636 dataset, which contains heterogeneous samples (background genetic variation, some batch effects). No further technical variation was added, and a foldchange was added on 300 genes.
* **seqc**: 5 vs 5 samples of seqc mixtures C and D respectively, which include two different spike-in mixes. The samples where selected from two batches with technical differences so that there is weak partial correlation with the mixtures (3:2 vs 2:3). Since the true differences between mixtures are not entirely known, the analysis was performed on all genes but only the spike-ins were considered for benchmarking.
* **simulation**: 8 vs 8 samples were simulated using `polyester` and the means/dispersions from the GSE79636 (restricted to a single batch and biological group), with 500 DEGs and two batch effects: 1) a technical batch partially correlated with the group (6:2) and affecting 1/3 of the genes, and 2) a linear vector of technical variation uncorrelated with the groups and affecting 1/3 of the genes.

@@

## Running the pipeline

```{r, eval=FALSE}
suppressPackageStartupMessages({
  library(SummarizedExperiment)
  library(pipeComp)
})
source(system.file("extdata", "dea_wrappers.R", package="pipeComp"))

pd <- dea_pipeline()

ds <- list.files("../data_preparation", pattern="\\.rds$", full.names = TRUE)
names(ds) <- gsub("\\.rds$","",basename(ds))

alternatives <- list(
  filt=c("none","def.filter"),
  minCount=c(10,20),
  sva.method=c("none","sva.svaseq","sva.vstsva","sva.RUVr","sva.RUVs"),
  k=c(1:5,"be","leek"),
  dea.method=c("dea.edgeR", "dea.edgeR.QLF", "dea.DESeq2", "dea.voom")
)

# we get rid of meaningless/redundant combinations:
eg <- buildCombMatrix(alternatives)
eg <- eg[eg$sva.method %in% c("sva.svaseq", "sva.vstsva") | 
           !(eg$k %in% c("be","leek")),]
eg <- eg[eg$filt=="def.filter" | eg$minCount==10,]
eg <- eg[eg$k=="1" | eg$sva.method!="none",]

res <- runPipeline(ds, alternatives, pd, comb = eg, saveEndResults = TRUE, debug=FALSE, nthreads = 3)
```

@@

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
})
theme_set(theme_cowplot())
devtools::load_all("../../pipeComp/")
res1b <- res1 <- res
res1$evaluation$dea <- lapply(res1$evaluation$dea, FUN=function(x) x[which(x$k==1),])
res1b$evaluation$dea <- lapply(res1b$evaluation$dea, FUN=function(x){
  x[x$filt=="def.filter" & x$minCount==10,]
})
```

```{r, fig.width=8, fig.height=9.5}
plot_grid(
  grid.grabExpr(draw(
    dea_evalPlot_logFCs(res1, agg.by=c("sva.method","dea.method"), reorder_rows=FALSE, row_split="sva.method") +
    dea_evalPlot_sig(res1, agg.by=c("sva.method","dea.method"), reorder_rows=FALSE)
  )),
  plotElapsed(res, agg.by=c("sva.method","dea.method")),
  nrow=2, labels=LETTERS[1:2]
)
```

```{r, fig.width=11, fig.height=9}
plot_grid(
  dea_evalPlot_curve(res1, agg.by=c("sva.method","dea.method"), colourBy="sva.method", shapeBy="dea.method") + ggtitle("Using a single surrogate variable"),
  dea_evalPlot_curve(res, agg.by=c("sva.method","k"), colourBy="k", shapeBy="sva.method") + ggtitle("Increasing the number of surrogate variables"),
  nrow=2, labels=LETTERS[1:2]
)
```

```{r, fig.width=12, fig.height=6.5}
plot_grid(
  plot_grid(
    dea_evalPlot_curve(res1, agg.by=c("dea.method","filt","minCount"), colourBy = "dea.method", shapeBy="filter") + ggtitle("DEA methods, Averaging across SVA strategies"),
    dea_evalPlot_curve(res1b, agg.by=c("sva.method")) + ggtitle("SVA methods, averaging across DEA methods"),
    nrow=2, labels=LETTERS[1:2]
  ),
  grid.grabExpr(draw(
    dea_evalPlot_sig(res, agg.by=c("sva.method","k"), reorder_rows=FALSE, 
                     show_column_names=TRUE, anno_legend = FALSE, row_split="sva.method")
    )),
  nrow=1, rel_widths=c(5,2), labels=c(NA,"C")
)
```

>>>>

@@

@@

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(pipeComp)
  library(grid)
  library(ComplexHeatmap)
})
theme_set(theme_cowplot())
res1b <- res1 <- res
res1$evaluation$dea <- lapply(res1$evaluation$dea, FUN=function(x) x[which(x$k==1),])
res1b$evaluation$dea <- lapply(res1b$evaluation$dea, FUN=function(x){
  x[x$filt=="def.filter" & x$minCount==10,]
})
```

```{r, fig.width=8, fig.height=9.5}
plot_grid(
  grid.grabExpr(draw(
    dea_evalPlot_sig(res1, what=c("logFC.pearson", "logFC.mad", "TPR", "FDR"), agg.by=c("sva.method","dea.method"), row_split="sva.method")
  )),
  plotElapsed(res, agg.by=c("sva.method","dea.method")),
  nrow=2, labels=LETTERS[1:2]
)
```

```{r, fig.width=12, fig.height=7}
plot_grid(
  plot_grid(
    dea_evalPlot_curve(res1, agg.by=c("sva.method","dea.method"), colourBy="dea.method", shapeBy="sva.method") + ggtitle("Using a single surrogate variable"),
    dea_evalPlot_curve(res, agg.by=c("sva.method","k"), colourBy="k", shapeBy="sva.method") + ggtitle("Increasing the number of surrogate variables"),
    nrow=2, labels=LETTERS[1:2]
  ),
  grid.grabExpr(draw(
    dea_evalPlot_sig(res, what=c("logFC.pearson", "TPR", "FDR"), agg.by=c("sva.method","k"), row_split="sva.method", 
                     anno_legend=FALSE, show_column_names=TRUE)
  )),
  nrow=1, rel_widths=c(5,2.5), labels=c(NA,"C")
)
```

```{r, fig.width=12, fig.height=6.5}
plot_grid(
  plot_grid(
    dea_evalPlot_curve(res1, agg.by=c("dea.method","filt","minCount"), colourBy = "dea.method", shapeBy="filter") + ggtitle("DEA methods, Averaging across SVA strategies"),
    dea_evalPlot_curve(res1b, agg.by=c("sva.method")) + ggtitle("SVA methods, averaging across DEA methods"),
    nrow=2, labels=LETTERS[1:2]
  ),
  grid.grabExpr(draw(
    dea_evalPlot_sig(res, what=c("TPR", "FDR"), agg.by=c("sva.method","k"), row_split="sva.method", 
                     anno_legend=FALSE, show_column_names=TRUE)
    )),
  nrow=1, rel_widths=c(5,2), labels=c(NA,"C")
)
```
